---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VFmixing

<!-- badges: start -->
<!-- badges: end -->

The goal of VFmixing is to provide methodology specifically designed to 
mix vector fields given noisy positional data.

## Installation

You can install the development version of VFmixing from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jrodu/VFmixing")
```

## Example

This is an example showing how to use the main set of functions. Let's look
at the noisy positional data from a particle moving through a some unknown 
vector field.

```{r}
library(VFmixing)

Data <- VFmixing::ExData
visualizeData(Data)
```

The key assumption for this method is that the true unknown vector field is a 
linear combination of a set of simpler known vector fields at every point in 
time. For this example, the set of mixing vector fields is a rotational field
and an expanding field. To use this set of vector fields in the mixing process
we construct the following function:

```{r}
baseVectorFields
```

Now that the baseVectorFields function has been defined, we can begin the
mixing process. This is broken into two steps: Path and Trajectory. The first
step involves estimating the path the particle took through the vector field.
We do this to get better estimates of both position and velocity that would
otherwise be compromised extremely quickly with only a small degree of error.
We provide three unique methods to estimate the path using cubic splines: 
"ByFours", "Full Regression", and "Projection". Each method has its pros and 
cons.

"ByFours":  
Pros: Extremely Fast Computation and Stable Estimates  
Con: No Estimable Covariance Structure
                
"Full Regression":  
Pros: Stable Estimates and a Estimable Covariance Structure  
Con: Slower Computation
                    
"Projection":  
Pros: Estimable Covariance Structure and Reasonably Quick Computation  
Con: Estimates may be unstable under extremely complex data
            
For complex real world scenarios, we recommend using the "ByFours" method, but 
for this example we will showcase the "Projection" method as the data is not too 
complex.

```{r}

# EstPath <- getEstimatedPaths_Projection(Data = Data, nsplits = 100, 
#                                         V_Smooth = F, 
#                                         Model_Selection_Type = "BIC")

EstPath <- VFmixing::ExEstPath_Projection
visualizeEstPath(Data = Data, EstPath = EstPath, t_grid_size = 0.01, 
                 x_var = 'X1', y_var = 'X2')

```

With the path of the particle estimated, we can move to the main event: the 
trajectory step. We want to estimate the weight functions for each 
mixing vector field. We call the weight functions the Trajectory of each
field. Again, we provide a couple unique methods to estimate the trajectory 
functions using cubic splines: "ByFours" and "One-Shot". Similar to the "ByFours" 
path method, this method runs extremely fast, outputs good estimates even 
with complex paths, but does not output a theoretical covariance structure. The 
"ByFours" method can be used regardless of which method was used to estimate the 
path. The "One-Shot" method uses the path covariance structure to transform 
the path coefficients into the trajectory coefficients. This method has a much 
higher computational cost and the estimates can sometimes falter with complex 
paths, but yields a theoretical covariance structure under the assumption 
that the mixing vector fields are relatively constant in a small window defined 
by the path error. The "One-Shot" method can only be used if the path was 
estimated using the "Full Regression" or "Projection" methods. In practice, we 
recommend using the "ByFours" method again, simply due to the much lower 
computational cost, but since we used the "Projection" methods for the path, we 
showcase the "One-Shot" method for the trajectories.

```{r}

# EstTraj <- getEstimatedTrajectories_OneShot(EstPath = EstPath, 
#                                                  nsplits = 100,
#                                      baseVectorFields = baseVectorFields,
#                                      V_Smooth = F,
#                                      Model_Selection_Type = "BICOpt",
#                                      Random_Path = F)

EstTraj <- VFmixing::ExEstTraj_OneShot
TrueTraj <- VFmixing::ExTraj
visualizeEstTraj(EstTraj = EstTraj, TrueTraj = TrueTraj$Traj, 
                 TrueTrajTimeSplits = TrueTraj$TimeSplits, EstPath = EstPath, 
                 baseVectorFields = baseVectorFields, t_grid_size = 0.01, 
                 CL = 0.95, P_Int = T, Cos_Cutoff = 0.91)
```

In the plot above, we can see our estimate of the trajectory (solid line)
next to the true trajectories used to construct the data (dashed line). 
Even with a reasonable amount of error in our data, visually, we did pretty well 
and the confidence region captures a good percentage of the true trajectory 
functions. However, we can use a couple more functions to evaluate our method's
performance using the average squared distance between the estimated 
trajectories and the truth, as well as calculating the coverage rate of the 
trajectories.

```{r}

AvgSqDist = cubicAvgSquareDist(EstCubic = EstTraj$Traj, 
                               TrueCubic = TrueTraj$Traj, 
                               EstTimeSplits = EstTraj$TimeSplits,
                               TrueTimeSplits = TrueTraj$TimeSplits, 
                               lower_bound = min(EstTraj$TimeSplits),
                               upper_bound = max(EstTraj$TimeSplits))
AvgSqDist

TrajCovRate = coverageRateTraj(EstTraj = EstTraj, TrueTraj = TrueTraj$Traj, 
                               TrueTrajTimeSplits = TrueTraj$TimeSplits, 
                               CL = 0.95)
TrajCovRate
```

We can see that the average squared distance is extremely small, but the 
coverage rate is lower than 95\%, which we would've expected. Therefore, the 
assumption that the mixing vector fields are constant within the error region
of a data point is not met, which we could have imagined due to the nature of the
vector fields and the amount of error in the data.
